Choosing TTC Subway Service Provides the Most Time Stability during Weekday Rush-Hours

### 2024-09-24 outputs/paper/paper.qmd
cut out several unused columns in cleaned delay data and re-rendered

```{r}
#| include: false
#| warning: false
#| message: false
# load librarys and datasets
library(tidyverse)
library(dplyr)
library(lubridate)
library(knitr)
library(ggplot2)
library(hms)
library(patchwork)

@@

# Data {#sec-data}
Where the data comes from

which datasets where chosen

why choose 2023

@@

knitr::kable(sample_bus_delay, col.names = c("Date", "Line", "Time", "Day", "Station", "Cause of Delay", "Min Delay", "Min Gap", "Direction", "Vehicle"), align = c("l","c", "c", "c", "c","c", "c", "c", "c","c"))
```

@@

# Generate the table with automatic text wrapping for column names
knitr::kable(sample_streetcar_delay, 
             col.names = c("Date", "Line", "Time", "Day", "Station", "Cause of Delay", "Min Delay", "Min Gap", "Bound", "Vehicle"), align = c("l","c", "c", "c", "c", "c", "c", "c", "c", "c")) 
```

@@

knitr::kable(sample_subway_delay, col.names = c("Date", "Time", "Day", "Station", "Delay Code", "Min Delayed", "Min Gap", "Bound", "Line", "Vehicle"), align = c("l","c", "c", "c", "c","c", "c", "c", "c","c"))
```

>>>>

```{r}
#| include: false
#| warning: false
#| message: false
# load librarys and datasets
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(hms)
library(patchwork)

@@

# Data {#sec-data}
Where the data comes from
- All data used for and throughout the paper are extracted from Open Data Toronto Portal [@OpenDataToronto]. TTC Bus Delay data [@Bus], TTC Streetcar Delay [@Streetcar], TTC Subway Delay [@Subway] and TTC Subway Delay Codes [@Subway] are retrieved to compare the delay magnitudes of different services. Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], dplyr [@dplyr], hms [@hms], ggplot2 [@ggplot2], patchwork [@patchwork], and knitr [@knitr].

For all delay dataset, only data in 2023 were chosen. There are three main reasons for this choice:

1. 2023 data is the most recent full-year data and offers the best reference for current (2024) delay patterns. 
2. COVID-19 impacted TTC operating capacity and styles in other years close to 2024. 
3. data from earlier years (before 2020) were structured very differently.

Therefore only 2023 delay data were chosen in this paper.

@@

knitr::kable(sample_bus_delay, col.names = c("Date", "Line", "Time", "Day", "Station", "Cause of Delay", "Min Delay"), align = c("l","c", "c", "c", "c","c", "c"))
```

@@

# Generate the table with automatic text wrapping for column names
knitr::kable(sample_streetcar_delay, 
             col.names = c("Date", "Line", "Time", "Day", "Station", "Cause of Delay", "Min Delay"), align = c("l","c", "c", "c", "c", "c", "c")) 
```

@@

knitr::kable(sample_subway_delay, col.names = c("Date", "Time", "Day", "Station", "Delay Code", "Min Delayed", "Line"), align = c("l","c", "c", "c", "c","c", "c"))
```
